---
title: "Lifetime incidence and age of onset of mental disorders, and 12-month service utilization in primary and secondary care: a Finnish nationwide registry study"
execute:
  echo: false
  warning: false
crossref:
  fig-prefix: Fig.
  ref-hyperlink: false
  custom:
    - kind: float
      reference-prefix: Supplementary Table
      key: supptbl
      space-before-numbering: true
    - kind: float
      reference-prefix: Supplementary Fig.
      key: suppfig
      space-before-numbering: true
    - kind: float
      reference-prefix: Extended Data Table
      key: exttbl
      space-before-numbering: true
    - kind: float
      reference-prefix: Extended Data Fig.
      key: extfig
      space-before-numbering: true
fig-dpi: 600
tbl-cap-location: top
editor: 
  markdown: 
    wrap: 72
format: html
filters:
  - docx-landscape.lua
---

```{r}
library(here)

# libraries, helper functions, ggplot theme, figure labels, diagnosis names: 
source(here("0_settings.R"), encoding = "UTF-8")

# read and process exported results from the Statistics Finland's closed analysis environment:
source(here("R", "01_read_process_data.R"), encoding = "UTF-8")
# draw figures:
source(here("R", "02_figs.R"), encoding = "UTF-8")
# draw supplementary figure:
source(here("R", "03_figs_washout.R"))

```


```{r}


make_tbl_cidg_data <- function(sample_in = "all", dgs_in = dgs_tbl[sub_chapter == T | dg %in% c("first", "first_no_f0"), dg]){
  dat <- tbl_cidg[sample == sample_in & dg %in% dgs_in] 
  
  diff <- dat[age == 100] %>% dcast(dg~sukup+age+sample, value.var = "est") %>%
  .[, w_to_m := get(paste0("Women_100_", sample_in)) / get(paste0("Men_100_", sample_in))]

  dat[, dg := factor(dg, levels = dgs_in)]

  dat <- rbindlist(
    list(
      dat,
      median_aoo[dg %in% dgs_in & sample == sample_in, .(sukup = factor(sukup, levels = c(2, 1), labels =  c("Women", "Men")), age = 101, sample, value, dg)]
      ),
    fill = T
  )
  
  dat %<>% dcast(dg~sukup+age+sample, value.var = "value") 

  dat <- dat[diff, on = "dg", w_to_m := i.w_to_m %>% numb(n_decim = 2)]

  dat %<>% 
  .[dg %in% dgs_in] %>%
  .[dgs_tbl, on = "dg", dg := i.labs] %>% 
  .[dg %in% dgs_tbl[subgr == T, labs], dg := paste0("€€", dg)] 
  
  setnames(dat, old = names(dat), new = names(dat) %>% str_remove(paste0("_", sample_in)))

  dat %<>% .[, !c("Women_80", "Men_80")] 
  dat[]
  
  # FILL NA 
  cols <- names(dat) %>%  setdiff("dg")
  dat[, (cols) := replace(.SD, is.na(.SD), "NA"), .SDcols = cols]
  
}

```


::: landscape
## Tables

### @tbl-cidg: Cumulative incidence of mental disorders at the ages of 25, 50, 75, and 100 years, and median age of onset (AOO) and interquartile range (IQR) by gender and ICD–10 sub-chapter category

::: {#tbl-cidg}
```{r}
make_tbl_cidg <- function(dat_in, no_indet = F, hl_subchapters = F){
  # A4 210 x 297 mm
  
  if(no_indet == TRUE){
    dat_in[, dg := dg %>% str_remove("€€")]
  }
  if(hl_subchapters == T){
    dat_in[dg %in% dgs_tbl[dg %in% c("first", paste0("f", seq(0,9))), labs], dg := paste0("**", dg, "**")]
  }
    
dat_in %>% 
  .[dg %in%  dgs_tbl[dg %in% c("first_no_f0"), labs], `:=`(dg = paste0("*", dg, "*"))] %>% 
    purrr::modify_if(is.factor, as.character) %>%
  flextable() %>% 
  set_header_labels(
    values = list(
      dg = labs_tbls$dg,
    Men_25 = labs_tbls$a25,
    Men_50 = labs_tbls$a50,
    Men_75 = labs_tbls$a75,
    Men_100 = labs_tbls$a100,
    Men_101 = labs_tbls$aoo,
    Women_25 = labs_tbls$a25,
    Women_50 = labs_tbls$a50,
    Women_75 = labs_tbls$a75,
    Women_100 = labs_tbls$a100,
    Women_101 = labs_tbls$aoo,
    w_to_m = ""
    )) %>% 
#  padding(i = ~ dg %in% dgs_tbl[subgr == T, labs], j= "dg", padding.left=20) %>% 
    add_header_row(values = c("", labs_tbls$women, labs_tbls$men, ""), colwidths = c(1, 5, 5, 1)) %>% 
    add_header_row(values = c("", labs_tbls$ci_title, paste0(labs_tbls$w_to_m, "^a^")), colwidths = c(1,10,1)) %>% 
  add_footer_lines(value = as_paragraph(as_sup("a"), labs_tbls$w_to_m_footnote)) %>% 
  theme_box() %>% 
    colformat_md(part = "all")
}

make_tbl_cidg_data() %>% make_tbl_cidg(no_indet = T)  %>% fit_table(landscaple = T)
```
:::
:::

\newpage

### @tbl-prevalAgegr: 12-month service utilization for medical contacts with diagnosed mental disorders by gender, age group, and type of contact in `r pr_year`^a^

::: {#tbl-prevalAgegr}
```{r}
tbl_preval_agegr %>% copy() %>% .[dg == "any"] %>% 
  #.[,agegr := factor(agegr, levels = c("0–24", "25–44", "45–64", "65–84", "85–99", "0–99"))] %>% 
  .[, sample := factor(sample, levels = labs_lst$sample$val, labels = labs_lst$sample$lab)] %>% 
  .[, sukup := factor(sukup, levels = labs_lst$sukup[order(ord1), val], labels = labs_lst$sukup[order(ord1), lab])] %>%
  .[, val := paste0(numb(n, n_decim = 0), " (", numb(pros), ")")] %>% 
  dcast(sample+sukup~agegr, value.var = "val") %>% 
  as_grouped_data(groups = "sample",) %>% 
  as_flextable(hide_grouplabel = TRUE) %>% 
  colformat_double(big.mark = numb_in$big_mark, decimal.mark = numb_in$decim_mark, digits = 2) %>% 
  autofit() %>% 
    theme_box() %>% 
  add_header_row(values = c("", labs_tbls$n_pr_title), colwidths = c(1, tbl_preval_agegr[, uniqueN(agegr)])) %>% 
  set_header_labels(
    values = list(
      sukup = "")) %>% 
  add_footer_lines(value = as_paragraph(as_sup("a"), labs_tbls$pr_footnote)) %>% 
  colformat_md(part = "all") %>% 
  fit_table() -> tbl

tbl
```
:::

\newpage

## Figures

### @fig-mainresults: Cumulative incidence, incidence rate, and 12-month service utilization^a^ of mental disorders by gender and treatment type

::: {#fig-mainresults}
```{r}
#| fig-width: 8
#| fig-height: 9.5

(fig_mr$a+fig_mr$b) /(fig_mr$c+fig_mr$d) / (fig_mr$e+fig_mr$f) +  plot_layout(guides = 'collect') & theme(legend.position = 'bottom')

```
:::

^a^ `r labs_tbls$pr_footnote`

\newpage

### @fig-dg1: Cumulative incidence, incidence rate, and 12-month service utilization^a^ of mental disorders by gender and diagnosis

```{r}
#| label: fig-dg1
#| fig-width: 9.1
#| fig-height: 8

fig_dgs$part1
```

^a^ `r labs_tbls$pr_footnote`

### @fig-dg2: Cumulative incidence, incidence rate, and 12-month service utilization^a^ of mental disorders by gender and diagnosis

```{r}
#| label: fig-dg2
#| fig-width: 9.1
#| fig-height: 8

fig_dgs$part2
```

^a^ `r labs_tbls$pr_footnote`

### @fig-dg3: Cumulative incidence, incidence rate, and 12-month service utilization^a^ of mental disorders by gender and diagnosis

```{r}
#| label: fig-dg3
#| fig-width: 9.1
#| fig-height: 8

fig_dgs$part3 
```

^a^ `r labs_tbls$pr_footnote`

\newpage

# Extended Data

::: landscape
### @exttbl-descriptive: Diagnostic classification of mental disorders according to the ICD–10, earliest possible age of onset, and the counts of cases and person-years included in the follow-up, and count of persons at risk at the start of the follow-up by gender

::: {#exttbl-descriptive}
```{r}
d <- datlist_incidence$descriptives$all %>% .[dg %in% dgs_tbl[specific_dgs == 1, dg] | dg %in% c("f_any", "f_any_no_f0", "f00_f03", "f04_f09")] %>% 
  .[dgs_tbl, on ="dg", c("min_age") := .(i.min_age)] %>% 
  .[, min_age := min_age %>% as.character()] %>% 
  .[dg %in% c("f_any", "f_any_no_f0"), `:=`(min_age = "NA")] %>% 
  .[dg %in% c("f_any", paste0("f", seq(0,9))), dg_lab := paste0("**", dg_lab, "**")] %>% 
  .[dg %in% c("f_any_no_f0"), `:=`(dg_lab = paste0("*", dg_lab, "*"))] %>% 
  .[, .(dg, dg_lab, min_age, Women_N, `Women_Person years`, Women_Diagnosis, Men_N, `Men_Person years`, Men_Diagnosis)]

  # FILL NA 
cols <- names(d)
d[, (cols) := replace(.SD, is.na(.SD), "NA"), .SDcols = cols]

d[, dg_lab := dg_lab %>% str_remove("€€")]

d[, -c("dg")] %>% flextable() %>% 
   set_header_labels(
    values = list(
      dg_lab = labs_tbls$dg,
      min_age = labs_tbls$min_age,
      Women_N = labs_tbls$n_indiv,
      `Women_Person years` = labs_tbls$n_py,
      Women_Diagnosis = labs_tbls$n_cases,
      Men_N = labs_tbls$n_indiv,
      `Men_Person years` = labs_tbls$n_py,
      Men_Diagnosis = labs_tbls$n_cases
    )) %>% 
      add_header_row(values = c("", labs_tbls$women, labs_tbls$men), colwidths = c(2, 3, 3)) %>% 
  #add_header_row(values = c("", labs_tbls$ci_title), colwidths = c(2,6)) %>% 
  theme_box() %>% 
  colformat_double(big.mark = numb_in$big_mark, decimal.mark = numb_in$decim_mark, digits = 0) %>% 
  colformat_md() %>% 
  fit_table(landscaple = T, supplement = T) -> tbl

tbl
```
:::

:::

### @extfig-washout: Example timelines in the main analysis and in sensitivity analyses with additional washout periods

::: {#extfig-washout}
```{r}
#| fig-width: 7
#| fig-height: 10

fig_washout()
```
:::

The procedures to differentiate incident cases from prevalent cases
demonstrated. Solid black lines indicate person-time in the study at
risk for truly incident cases. Red lines indicate person-time in the
study where participants had previous treatments before the start of the
washout period and were incorrectly included. Dashed lines indicate
person-time that is not included in the study. Complete washout
indicates the washout period including inpatient, outpatient and primary
care treatments simultaneously.

**Panel A** demonstrates the main analysis, in which we excluded
disorder-specific prevalent cases at the start of the follow-up period.
This exclusion included individuals with inpatient treatments between
1975 and 1999, as well as those with outpatient care between 1998 and
1999.

**Panel B** demonstrates the sensitivity analysis with an additional
three-year washout period and a retrospective washout for primary care.
In this analysis, we shortened the study period from
`r tms$start`–`r tms$end` to `r tms$sens1_start`–`r tms$end`,
incorporating an extra one-year washout for primary care. This approach
aligns the washout procedure with that of a previous Danish study, where
medication use served as a proxy for primary care
contacts[@kessing2023]. While this method is stricter in excluding
prevalent cases, it introduces immortal time bias. In the Danish study,
estimates of lifetime risk for mental disorders with additional 5- and
10-year washout periods were reported, with the estimates decreasing by
only 0.6 and 2.0 percentage points, respectively.

**Panel C** demonstrates the sensitivity analysis with a further
restricted follow-up period. In this analysis, we shortened the study
period from`r tms$start`–`r tms$end` to `r tms$sens2_start`–`r tms$end`.
The washout period for inpatient care is
`r tms$inpat`–`r tms$sens2_start-1`, for outpatient secondary care
`r tms$outpat`–`r tms$sens2_start-1`, and for primary care the year
`r tms$prim`. This scenario avoids introducing immortal time bias but
limits the overall time span captured in the study.

The following example timelines are presented:

Person 1 has his/her first inpatient admission during the study period,
with no prior outpatient or primary care mental health treatments, and
is therefore correctly identified as an incident case.

Person 2 has his/her first psychiatric outpatient contact during the
study period, with no prior inpatient or primary care mental health
treatments, and is likewise correctly identified as an incident case.

Person 3 has his/her first contact with primary care mental health
services in `r tms$prim` or later, with no prior inpatient or secondary
care mental health treatments, and is thus correctly identified as an
incident case.

Person 4 contributes to the risk of experiencing a mental disorder until
he/she dies or emigrates from Finland.

Person 5 contributes to the risk of experiencing a mental disorder until
the end of the study period.

Person 6 had his/her first inpatient admission before the launch of
inpatient data in the register (`r tms$inpat`). Consequently, he/she
enters the study incorrectly and contributes to the risk of an incident
mental disorder. In this case, he/she has an inpatient admission during
the study period that is misclassified as an incident event.

Person 7 had his/her first inpatient admission before the launch of
inpatient data in the register (`r tms$inpat`) and a psychiatric
secondary care outpatient treatment before the launch of outpatient data
in the register (`r tms$outpat`). Thus, he/she enters the study
incorrectly and contributes to the risk of an incident mental disorder.
In this example, he/she has an outpatient contact during the study
period that is misclassified as an incident event.

Person 8 had his/her first inpatient admission before the launch of
inpatient data in the register (`r tms$inpat`), a psychiatric secondary
care outpatient treatment before the launch of outpatient data in the
register (`r tms$outpat`), and a primary care mental health contact
before the launch of primary care data in the register (`r tms$prim`).
Therefore, he/she enters the study incorrectly and contributes to the
risk of an incident mental disorder. In this case, he/she has a primary
care mental health contact in `r tms$prim` or later and is misclassified
as an incident event.

Persons 9 and 10 are examples of individuals who had inpatient care
between `r tms$inpat` and `r tms$start -1` or secondary outpatient care
between `r tms$outpat` and `r tms$start -1`, and are correctly excluded
from the study population.

Person 11 has his/her first contact with primary care mental health
services in `r tms$prim` and demonstrates the additional primary care
washout in the sensitivity analyses.

# Supplementary Information

\newpage

::: landscape
### @supptbl-cidgESH: Cumulative incidence of mental disorders in psychiatric secondary care at the ages of 25, 50, 75, and 100 years, and median age of onset (AOO) and interquartile range (IQR) by gender and ICD–10 sub-chapter category

::: {#supptbl-cidgESH}
```{r}
make_tbl_cidg_data(sample_in = "esh") %>% make_tbl_cidg(no_indet = T) %>% fit_table(landscaple = T, supplement = T)
```
:::

:::

::: landscape
### @supptbl-cidgINPAT: Cumulative incidence of diagnosed mental disorders in psychiatric inpatient care at the ages of 25, 50, 75, and 100 years, and median age of onset (AOO) and interquartile range (IQR) by gender and ICD–10 sub-chapter category

::: {#supptbl-cidgINPAT}
```{r}
make_tbl_cidg_data(sample_in = "inpat") %>% make_tbl_cidg(no_indet = T) %>% fit_table(landscaple = T, supplement =  T)
```
:::

:::

::: landscape
### @supptbl-cidg-alldgs: Cumulative incidence of mental disorders at the ages of 25, 50, 75, and 100 years, and median age of onset (AOO) and interquartile range (IQR) by gender and diagnosis

::: {#supptbl-cidg-alldgs}
```{r}
make_tbl_cidg_data(dgs_in = dgs_tbl[specific_dgs == 1, dg]) %>% make_tbl_cidg(no_indet = T, hl_subchapters = T)  %>% fit_table(landscaple = T, supplement = T)
```
:::

:::

### @suppfig-cicomparedg: Cumulative incidence of mental disorders by gender and ICD–10 sub-chapter category

::: {#suppfig-cicomparedg}
```{r}
#| fig-width: 7.5
#| fig-height: 7

fig_compare_dgs$ci
```
:::

\newpage

::: landscape
### @supptbl-prevalAgegr-alldg: 12-month service utilization for medical contacts with diagnosed mental disorders by gender, age group, and diagnosis in `r pr_year`^a^

::: {#supptbl-prevalAgegr-alldg}
```{r, missa}
d <- tbl_preval_agegr %>% copy() %>% .[sample == "all"] %>% 
  #.[,agegr := factor(agegr, levels = c("0–24", "25–44", "45–64", "65–84", "85–99", "0–99"))] %>% 
  .[, sample := factor(sample, levels = labs_lst$sample$val, labels = labs_lst$sample$lab)] %>% 
  .[, sukup := factor(sukup, levels = labs_lst$sukup[order(ord1), val], labels = labs_lst$sukup[order(ord1), lab])] %>%
  .[, val := paste0(numb(n, n_decim = 0), " (", numb(pros), ")")] %>% 
  dcast(dg~sukup+agegr, value.var = "val")

d %<>% 
#  .[dg %in% dgs_in] %>%
  .[dgs_tbl, on = "dg", dg := i.labs] %>% 
  .[dg %in% dgs_tbl[subgr == T, labs], dg := paste0("€€", dg)] 

d[dg %in%  dgs_tbl[dg %in% c("any_no_f0"), labs], `:=`(dg = paste0("*", dg, "*"))]

d[, dg := dg %>% str_remove("€€")]
d[dg %in% dgs_tbl[dg %in% c("first", paste0("f", seq(0,9))), labs], dg := paste0("**", dg, "**")]  


d %>% flextable() %>% 
  colformat_double(big.mark = numb_in$big_mark, decimal.mark = numb_in$decim_mark, digits = 2) %>% 
  set_header_df(mapping = data.frame(keys = d %>% names(), values = d %>% names() %>% word(2, sep = "_"), stringsAsFactors = FALSE), key = "keys" ) %>% 
  autofit() %>% 
  theme_box() %>%  
  add_header_row(values = c("", labs_tbls$women, labs_tbls$men), colwidths = c(1, 5, 5)) %>% 
  add_header_row(values = c("", labs_tbls$n_pr_title), colwidths = c(1, tbl_preval_agegr[, uniqueN(agegr)]*2)) %>% 
    add_footer_lines(value = as_paragraph(as_sup("a"), labs_tbls$pr_footnote)) %>% 
  colformat_md() %>% 
  fit_table(landscaple = T, supplement = T) -> tbl

tbl
```
:::

:::
